Mechanistic Modeling and Analysis of the Mammalian Unfolded Protein Response

### 2016-07-18 scientific-reports-title/Paper_v2.tex
Lina's edits

\begin{abstract}
Cells monitor protein folding by an inbuilt quality-control system in which incorrectly or misfolded proteins are tagged for degradation or sent back through a refolding cycle.
However, continued accumulation of incorrectly folded proteins triggers the Unfolded Protein Response (UPR), which attempts to re-establish folding homeostasis or commits the cell to apoptosis.
In this study, we developed a family of mechanistic models of the mammalian UPR system.
An ensemble of models parameters was estimated by minimizing the difference between simulations and experimental measurements using multiobjective optimization.
The ensemble of model parameters was validated using cross-validation.
Analysis of the model ensemble suggested the three branches of UPR fired simultaneously.
However, the importance of each branch was rank ordered in time; PERK and IRE1 were more important early, while ATF6 was important later in the response.
The activity of all three branches was coordinated by the molecular chaperone BiP.
Model analysis suggested that BiP feedback was critical to the overall robustness of the system.
Removal of any one branch of BiP feedback, destabilized the other branches.
On the other hand, removal of all nodes of BiP feedback increased the overall robustness of the system.
Thus, while BiP feedback is crucial to allowing the cell to adapt to small perturbations, it also makes the system fragile and susceptible to manipulation.

>>>>

\begin{abstract}
Cells monitor protein folding by an inbuilt quality-control system in which incorrectly folded (or misfolded) proteins are tagged for degradation or sent back through a refolding cycle.
However, continued accumulation of incorrectly folded proteins triggers the Unfolded Protein Response (UPR), which attempts to re-establish folding homeostasis or commits the cell to apoptosis.
In this study, we developed a family of mechanistic models of the mammalian UPR system.
An ensemble of models parameters was estimated by minimizing the difference between simulations and experimental measurements using multiobjective optimization.
The ensemble of model parameters was validated using cross-validation.
Analysis of the model ensemble suggested the three branches of UPR fired simultaneously.
However, the importance of each branch was rank ordered in time; PERK and IRE1 were more important early, while ATF6 was important later in the response.
The activity of all three branches was coordinated by the molecular chaperone BiP.
Model analysis suggested that BiP feedback was critical to the overall robustness of the system.
Removal of any one branch of BiP feedback, destabilized the other branches.
On the other hand, removal of all nodes of BiP feedback increased the overall robustness of the system.
Thus, while BiP feedback is crucial to allowing the cell to adapt to small perturbations, it also makes the system fragile and susceptible to manipulation.
